FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-06 ess_marmo.tex
Update on Overleaf.

 \authors{C. Marmo\affil{1},
 T. M. Hare\affil{2},
 S. Erard\affil{3},
 B. Cecconi\affil{3},
 M. Minin\affil{4},
 A. P. Rossi\affil{4} ...}

@@

FITS is already able to propose standard formatting for some data products
quite common by now in planetary surface investigations.
In particular, the Multi-Extension FITS (MEF) schema proposes an easy way
to store inhomogeneous digital information (reflectance, calibration data,
vector, table data, etc.) in the same file, each with correspondent metadata,
as well as multi-detector imagery (e.g. from HiRISE \citep{hirise})
or hyperspectral cubes with geometry information (e.g. from CRISM
\citep{crism} or OMEGA \citep{omega} instruments).
FITS has been already chosen to distribute, e.g.,
Hayabusa\footnote{http://darts.isas.jaxa.jp/planet/project/hayabusa/index.html},
Akatsuki\footnote{http://darts.isas.jaxa.jp/planet/project/akatsuki/index.html}
and some of the Dawn\footnote{https://sbn.psi.edu/pds/archive/dawn.html} data.
To be efficiently used in planetary surface investigations, FITS metadata must
be extended in order to take into account the size of the reference body.
Planetary body shapes and orientations are well defined and standardized by
the IAU Working Group on Cartographic Coordinates and Rotational Elements
(WGCCRE).
This group reports triennially \citep{wgccre} on the preferred rotation rate,
spin axis, prime meridian, and reference surface for planets and satellites
ensuring that cartographic endeavors are effectively comparable.
In the framework of the VESPA\footnote{http://europlanet-vespa.eu/}
\citep{vespapss} workpackage of the EuroplanetRI2020\footnote{http://www.europlanet-2020-ri.eu/}
project an extension to FITS metadata (GeoFITS) has been
proposed\footnote{https://voparis-confluence.obspm.fr/display/VES/GeoFITS:+Planetary+Data+FITS+format+and+metadata+convention},
taking into account the WGCCRE recommendations.

@@

Multiple coordinate system descriptions are allowed in the same header: this is a valuable
property for planetary surface imagery, as angular ('deg') and linear ('m') description of
the surface are often needed at the same time.
Further below 3D keyword will be proposed for FITS metadata, in order to standardize
the conversion between angular and linear coordinates: alternate description will still
be important in our framework as standard libraries (e.g.
wcslib\footnote{http://www.atnf.csiro.au/people/mcalabre/WCS/wcslib/} or
astropy\footnote{http://www.astropy.org/} \citet{astropy}) and some FITS visualization
software (e.g. ds9\footnote{http://ds9.si.edu/}) already implement them.

@@

\subsection{Examples and Best Practices}
\subsubsection{Global maps}
\subsubsection{HiRISE mosaics}
\subsubsection{OMEGA spectra and geometry data}
\subsubsection{Akatsuki imaging data}
\section{Inter-operable software developments}
\label{sec:softdev}

>>>>

 \authors{C. Marmo\affil{1},
 T. M. Hare\affil{2},
 S. Erard\affil{3},
 B. Cecconi\affil{3},
 M. Minin\affil{4},
 A. Zinzi\affil{5},
 A. P. Rossi\affil{4}}

@@

FITS is already able to propose standard formatting for some data products
quite common by now in planetary surface investigations.
In particular, the Multi-Extension FITS (MEF) schema proposes an easy way
to store inhomogeneous digital information (reflectance, calibration data,
vector, table data, etc.) in the same file, each with correspondent metadata,
as well as multi-detector imagery (e.g. from HiRISE \citep{hirise})
or hyperspectral cubes with geometry information (e.g. from CRISM
\citep{crism} or OMEGA \citep{omega} instruments).
FITS has been already chosen to distribute, e.g.,
Hayabusa\footnote{http://darts.isas.jaxa.jp/planet/project/hayabusa/index.html},
Akatsuki\footnote{http://darts.isas.jaxa.jp/planet/project/akatsuki/index.html}
and some of the Dawn\footnote{https://sbn.psi.edu/pds/archive/dawn.html} data.
To be efficiently used in planetary surface investigations, FITS metadata must
be extended in order to take into account the size of the reference body.
Planetary body shapes and orientations are well defined and standardized by
the IAU Working Group on Cartographic Coordinates and Rotational Elements
(WGCCRE).
This group reports triennially \citep{wgccre} on the preferred rotation rate,
spin axis, prime meridian, and reference surface for planets and satellites
ensuring that cartographic endeavors are effectively comparable.
In the framework of the VESPA\footnote{http://europlanet-vespa.eu/}
\citep{vespapss} workpackage of the EuroplanetRI2020\footnote{http://www.europlanet-2020-ri.eu/}
project an extension to FITS metadata (GeoFITS) has been
proposed\footnote{https://voparis-confluence.obspm.fr/display/VES/GeoFITS:+Planetary+Data+FITS+format+and+metadata+convention},
taking into account the WGCCRE recommendations \citep{MarmoLPSC}.

@@

Multiple coordinate system descriptions are allowed in the same header: this is a valuable
property for planetary surface imagery, as angular ('deg') and linear ('m') description of
the surface are often needed at the same time.
Further below 3D keyword will be proposed for FITS metadata, in order to standardize
the conversion between angular and linear coordinates: alternate description will still
be important in our framework as standard libraries (e.g.
wcslib\footnote{http://www.atnf.csiro.au/people/mcalabre/WCS/wcslib/} or
astropy\footnote{http://www.astropy.org/} \citet{astropy}) and some FITS visualization
software (e.g. ds9\footnote{http://ds9.si.edu/}) already implement them, making linear
coordinate tracking more efficient.

@@

\subsection{Examples and Best Practices}

\subsubsection{Visualization of global and projected maps}
High level products are meant to be used for visual inspection or GIS analysis.
They are distributed as projected rasters, often in GeoTIFF\footnote{http://geotiff.osgeo.org/}
format, as GeoTIFF make them GIS ready.
Once new keywords introduced in FITS and WCS definitions adapted to planetary surface, FITS
and GeoTIFF are completely interchangeable.
As an example we have downloaded some global maps from USGS
Astropedia\footnote{https://astrogeology.usgs.gov/search} and converted them in FITS.
We used ds9 to display the FITS image and were able to visualize coordinates in meters
as in QGIS, but also in planetary latitude and longitude thanks to alternate coordinate
description.

\subsubsection{HiRISE mosaics}

\subsubsection{OMEGA spectra and geometry data}
\subsubsection{Akatsuki imaging data}
\section{Inter-operable software developments}
\label{sec:softdev}
